---
title: "Regression -- relationship between asthma and other covariates"
author: "Xuewei Zhang"
data: "12/08/2017"
output: html_document

---

##BST 260 Final Project
#Group: Chuying Ma, Jingjing Tang, Jie Yin, Xuewei Zhang

#This Rmd file is for plotting regression for asthma and other covariates of interests

```{r}
###read data with exposure: greenness, pm2.5, ozone
library(dslabs)
library(tidyverse)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)

year_expo3 = read.csv("data_3_expo.csv")
names(year_expo3)
year_expo4 = read.csv("data_exposure_00_10.csv")
```


relationship between asthma and smoke
```{r}
year_expo3 %>%
  dplyr::select(-c(fips,state,county)) %>%
  ggplot(aes(asthma,smoke)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ year) +
  ggtitle("heartattack vs ozone")

```

##Conclusion
Not very clear univariate relationship

relationship between asthma and income1, which has a relatively positively relationship
```{r}
year_expo4 %>%
  dplyr:: select(-c(fips,state,county)) %>%
  ggplot(aes(asthma,income1)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ year) +
  scale_x_continuous(limits=c(0, 0.5)) +
  scale_y_continuous(limits=c(0, 0.5))
```

##Conclusion
slightly positive relationship between asthma and the proportional of people in the low income group

