---
title: "Classification -- applying LDA & QDA"
author: "Xuewei Zhang"
data: "12/08/2017"
output: html_document


---

##BST 260 Final Project
##Group: Chuying Ma, Jingjing Tang, Jie Yin, Xuewei Zhang

This Rmd file is for useing LDA and QDA to predict class of mental health

```{r}
library(dplyr)
library(ggplot2)
library(ISLR)
library(MASS)
library(mvtnorm)
```

##Part 1. LDA

read data of train and test
```{r,warning=FALSE}
train_lda = read.csv("train.csv")
test_lda = read.csv("test.csv")
attach(train_lda)
attach(test_lda)

names(train_lda)
names(test_lda)


###Use LDA to predict data and compare them with test data

lda.fit=lda(mental_class~.,data=train_lda)
lda.fit
lda.pred=predict(lda.fit, test_lda)
lda.class=lda.pred$class 
table(lda.class,test_lda$mental_class)
(390+1094)/(159+319+390+1094) 
####picture!
ldahist(data = lda.pred$x[,1], g=mental_class)
#####

###picture for two exp:
lda.fit2=lda(mental_class~student+employed,data=train_lda)
lda.fit2
lda.pred2=predict(lda.fit2, test_lda)
ldahist(data = lda.pred2$x[,1], g=mental_class)

train_lda = train_lda[complete.cases(train_lda),]
test_lda = test_lda[complete.cases(test_lda),]


train_lda %>% mutate(mental_class = factor(mental_class)) %>% 
  ggplot(aes(employed, student, fill = mental_class, color=mental_class)) + 
  geom_point(pch=21,cex=5, color="black") + 
  stat_ellipse(lwd=2, type="norm")

```

##Conclusion
the accuracy rate for LDA is around 0.7564, which is not bad

##Part 2. QDA

Use QDA to predict data and compare them with test data
```{r}
qda.fit=qda(mental_class~.,data=train_lda)
qda.fit
qda.pred=predict(qda.fit, test_lda)
qda.class=qda.pred$class 
table(qda.class,test_lda$mental_class)
(360+1076)/(349+177+360+1076) 

```

##Conclusion
the accuracy rate is about 0.7319, which is slightly lower than LDA