---
title: "Mental_part2"
author: "Chuying Ma"
date: "11/28/2017"
output: html_document
---

BST 260 Final Project
Group: Chuying Ma, Jingjing Tang, Jie Yin, Xuewei Zhang

This Rmd file is for conducting linear regression for menthlth and other covariates.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(tidyverse)
```

look for regression with pm 2.5, using backward selection and choose p = 0.05 as criteria:
```{r}
data_all = read_csv("data_3_expo.csv")
reg_data = data_all %>%
  dplyr::select(-c(fips,state,state,year))
```

```{r}
pm_fit1 = lm(menthlth ~ .,data = reg_data)
summary(pm_fit1)
```

delete age:
```{r}
pm_fit2 = lm(menthlth ~ sex + white + black + asian + educ1 + educ2 + income1 + income2 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data)
summary(pm_fit2)
```
delete white:
```{r}
pm_fit3 = lm(lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income2 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit3)
```
delete income2:
```{r}
pm_fit4 = lm(lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit4)
```

delete income3:
```{r}
pm_fit5 = lm(lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit5)
```

delete genhlth_vg:
```{r}
pm_fit6 = lm(lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex  + genhlth_go + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit6)
```

delete genhlth_go:
```{r}
pm_fit7 = lm(lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit7)
```

delete educ1:
```{r}
pm_fit8 = lm(lm(menthlth ~ sex + black + asian + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit8)
```

delete bmi_cts:
```{r}
pm_fit9 = lm(lm(menthlth ~ sex + black + asian + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit9)
```

delete income5:
```{r}
pm_fit10 = lm(lm(menthlth ~ sex + black + asian + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit10)
```

delete income7:
```{r}
pm_fit11 = lm(lm(menthlth ~ sex + black + asian + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit11)
```

delete asian:
```{r}
pm_fit12 = lm(lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit12)
```

delete bmi_cat1:
```{r}
pm_fit13 = lm(lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit13)
```

delete bmi_cat2:
```{r}
pm_fit14 = lm(lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit14)
```

delete income6:
```{r}
pm_fit15 = lm(lm(menthlth ~ sex + black + educ2 + income1 + income4 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit15)
```

delete exercise:
```{r}
pm_fit16 = lm(lm(menthlth ~ sex + black + educ2 + income1 + income4 + married + employed + out_of_work + homemaker + student + hlthplan + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit16)
```

delete sex:
```{r}
pm_fit17 = lm(lm(menthlth ~ black + educ2 + income1 + income4 + married + employed + out_of_work + homemaker + student + hlthplan + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + pm2.5,data = reg_data))
summary(pm_fit17)
```

look for regression with ozone, using backward selection and p = 0.05 as criteria:
```{r}
reg_data = data_all %>%
  select(-c(fips,state,county,year,heartattack,ACheartdis,stroke,asthma,pm2.5,greenness))
```

```{r}
oz_fit1 = lm(menthlth ~ .,data = reg_data)
summary(oz_fit1)
```

delete white:
```{r}
oz_fit2 = lm(menthlth ~ age + sex + black + asian + educ1 + educ2 + income1 + income2 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit2)
```

delete age:
```{r}
oz_fit3 = lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income2 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit3)
```

delete genhlth_vg:
```{r}
oz_fit4 = lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income2 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit4)
```

delete genhlth_go:
```{r}
oz_fit5 = lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income2 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit5)
```

delete income2:
```{r}
oz_fit6 = lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit6)
```

delete income3:
```{r}
oz_fit7 = lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit7)
```

delete bmi_cts:
```{r}
oz_fit8 = lm(menthlth ~ sex + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit8)
```

delete educ1:
```{r}
oz_fit9 = lm(menthlth ~ sex + black + asian + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit9)
```

delete income5:
```{r}
oz_fit10 = lm(menthlth ~ sex + black + asian + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit10)
```


delete asian:
```{r}
oz_fit11 = lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit11)
```


delete bmi_cat1:
```{r}
oz_fit12 = lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat2 + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit12)
```

delete bmi_cat2:
```{r}
oz_fit13 = lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit13)
```

delete income7:
```{r}
oz_fit14 = lm(menthlth ~ sex + black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit14)
```

delete sex:
```{r}
oz_fit15 = lm(menthlth ~ black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit15)
```

delete income6:
```{r}
oz_fit16 = lm(menthlth ~ black + educ2 + income1 + income4 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit16)
```

delete exercise:
```{r}
oz_fit17 = lm(menthlth ~ black + educ2 + income1 + income4 + married + employed + out_of_work + homemaker + student + hlthplan + smoke + drink + genhlth_ex + genhlth_fa + genhlth_po + qlactlm + ozone,data = reg_data)
summary(oz_fit17)
```

look for regression with greenness, using backward selection and p = 0.05 as criteria:
```{r}
reg_data = data_all %>%
  select(-c(fips,state,county,year,heartattack,ACheartdis,stroke,asthma,pm2.5,ozone))
```

```{r}
gr_fit1 = lm(menthlth ~ .,data = reg_data)
summary(gr_fit1)
```

delete income2:
```{r}
gr_fit2 = lm(menthlth ~ age + sex + white + black + asian + educ1 + educ2 + income1 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit2)
```

delete white:
```{r}
gr_fit3 = lm(menthlth ~ age + sex  + black + asian + educ1 + educ2 + income1 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit3)
```

delete age:
```{r}
gr_fit4 = lm(menthlth ~ sex  + black + asian + educ1 + educ2 + income1 + income3 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit4)
```

delete income3:
```{r}
gr_fit5 = lm(menthlth ~ sex  + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_vg + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit5)
```

delete genhlth_vg:
```{r}
gr_fit6 = lm(menthlth ~ sex  + black + asian + educ1 + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit6)
```

delete educ1:
```{r}
gr_fit7 = lm(menthlth ~ sex  + black + asian + educ2 + income1 + income4 + income5 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit7)
```

delete income5 :
```{r}
gr_fit8 = lm(menthlth ~ sex  + black + asian + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + bmi_cts + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit8)
```

delete bmi_cts:
```{r}
gr_fit9 = lm(menthlth ~ sex  + black + asian + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat1 + bmi_cat2 + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit9)
```

delete bmi_cat1:
```{r}
gr_fit10 = lm(menthlth ~ sex  + black + asian + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + bmi_cat2 + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit10)
```

delete bmi_cat2:
```{r}
gr_fit11 = lm(menthlth ~ sex  + black + asian + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit11)
```

delete asian:
```{r}
gr_fit12 = lm(menthlth ~ sex  + black + educ2 + income1 + income4 + income6 + income7 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit12)
```

delete income7:
```{r}
gr_fit13 = lm(menthlth ~ sex  + black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit13)
```

delete sex:
```{r}
gr_fit14 = lm(menthlth ~ black + educ2 + income1 + income4 + income6 + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit14)
```

delete income6:
```{r}
gr_fit15 = lm(menthlth ~ black + educ2 + income1 + income4  + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex + genhlth_go + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit15)
```

delete genhlth_go:
```{r}
gr_fit16 = lm(menthlth ~ black + educ2 + income1 + income4  + married + employed + out_of_work + homemaker + student + hlthplan + exercise + smoke + drink + genhlth_ex  + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit16)
```

delete homemaker:
```{r}
gr_fit17 = lm(menthlth ~ black + educ2 + income1 + income4  + married + employed + out_of_work  + student + hlthplan + exercise + smoke + drink + genhlth_ex  + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit17)
```

delete student:
```{r}
gr_fit18 = lm(menthlth ~ black + educ2 + income1 + income4  + married + employed + out_of_work + hlthplan + exercise + smoke + drink + genhlth_ex  + genhlth_fa + genhlth_po + qlactlm + greenness,data = reg_data)
summary(gr_fit18)
```


##conclusion

Taking mental health as example, multiple regression was applied to address the impact of environmental exposures adjusting for other socioeconomic factors, which shows a significant relationship between mental health and other environmental exposures and socioeconomic variables