---
title: "Visualization Graphs"
author: "Jingjing_Tang"
date: "11/27/2017"
output: html_document
---

BST 260 Final Project
Group: Chuying Ma, Jingjing Tang, Jie Yin, Xuewei Zhang

This Rmd file is for data analysis and visualization.

```{r}
library(maps)
library(ggplot2)
library("SASxport")
library(ggmap)
library(mapdata)
library(RColorBrewer)
library(gganimate)
library(dplyr)
dat <- read.csv("data_3_expo.csv")
dat051<- read.csv("state_05.csv") %>% select(-full, -X)
dat061<- read.csv("state_06.csv") %>% select(-full, -X)
dat071<- read.csv("state_07.csv") %>% select(-full, -X)
dat081<- read.csv("2008_data_state.csv") 
dat091<- read.csv("2009_data_state.csv") 
dat101<- read.csv("2010_data_state.csv") 
dat_state_05_10 <- rbind (dat051, dat061, dat071, dat081,dat091,dat101)
```


Our rough guesses according to the spatial distributions is confirmed by the correlation information among diseases.

Mental status and heart diseases are strongly related while the correlations between asthma and other disease are very small.

Based on this analysis, we use the information of heart attack to represent the three types of heart diseases. 

#correlation among diseases

```{r}
library(GGally)
 
# Check correlation between variables
cor(dat_state_05_10[3:7]) 
 
# Check correlations (as scatterplots), distribution and print corrleation coefficient 
ggpairs(dat_state_05_10[3:7]) 
 
# Nice visualization of correlations
ggcorr(dat_state_05_10[3:7], method = c("everything", "pearson")) 
```



#expo VS income/race/edu
```{r}
library(tidyverse)
dat_raw08 <- read.csv("2008_data_raw.csv") %>% mutate(race = ORACE2, educ = EDUCA, income = INCOME2, menthlth = MENTHLTH, heartattack = CVDINFR4, ACheartdis = CVDCRHD4, stroke = CVDSTRK3, asthma = ASTHMA2) %>%
  select(race, educ, income, menthlth, heartattack, ACheartdis, stroke, asthma) 
dat_raw09 <- read.csv("2009_data_raw.csv") %>% mutate(race = ORACE2, educ = EDUCA, income = INCOME2, menthlth = MENTHLTH, heartattack = CVDINFR4, ACheartdis = CVDCRHD4, stroke = CVDSTRK3, asthma = ASTHMA2) %>%
  select(race, educ, income, menthlth, heartattack, ACheartdis, stroke, asthma) 
dat_raw010 <- read.csv("2010_data_raw.csv") %>% mutate(race = ORACE2, educ = EDUCA, income = INCOME2, menthlth = MENTHLTH, heartattack = CVDINFR4, ACheartdis = CVDCRHD4, stroke = CVDSTRK3, asthma = ASTHMA2) %>%
  select(race, educ, income, menthlth, heartattack, ACheartdis, stroke, asthma) 
dat_raw <- rbind(na.omit(dat_raw08), na.omit(dat_raw09), na.omit(dat_raw010)) %>% mutate(group = c(NA)) 
dat_raw$educ[which(dat_raw$educ == 1 | dat_raw$educ == 2)] <- c("value1")
dat_raw$educ[which(dat_raw$educ == 3 | dat_raw$educ == 4)] <- c("value2")
dat_raw$educ[which(dat_raw$educ == 5 | dat_raw$educ == 6)] <- c("value3")
dat_raw$income[which(dat_raw$income == 1 | dat_raw$income == 2)] <- c("value1")
dat_raw$income[which(dat_raw$income == 3 | dat_raw$income == 4)] <- c("value2")
dat_raw$income[which(dat_raw$income == 5 | dat_raw$income == 6)] <- c("value3")
dat_raw$income[which(dat_raw$income == 7 | dat_raw$income == 8)] <- c("value4")
dat_raw$race[which(dat_raw$race == 1 )] <- c("value1")
dat_raw$race[which(dat_raw$race == 2 )] <- c("value2")
dat_raw$race[which(dat_raw$race == 3 )] <- c("value3")
dat_raw$race[which(dat_raw$race == 4 | dat_raw$race == 5 |dat_raw$race == 6)] <- c("value4")
dat_raw <- dat_raw %>% group_by(income) %>% mutate(income_num = n()) %>% ungroup() %>%
  group_by(educ) %>% mutate(educ_num = n()) %>% ungroup() %>%
  group_by(race) %>% mutate(race_num = n()) %>% ungroup()


dat_raw_menthlth <- dat_raw %>% filter(menthlth > 10) %>% mutate(group = c("A"))
dat_raw_heartattack <- dat_raw %>% filter(heartattack == 1) %>% mutate(group = c("B"))
dat_raw_ACheartdis <- dat_raw %>% filter(ACheartdis == 1) %>% mutate(group = c("C"))
dat_raw_stroke <- dat_raw %>% filter(stroke ==1) %>% mutate(group = c("D"))
dat_raw_asthma <- dat_raw %>% filter(asthma ==1) %>% mutate(group = c("E"))
dat_raw_com <- rbind(dat_raw_menthlth, dat_raw_heartattack, dat_raw_ACheartdis, dat_raw_stroke, dat_raw_asthma)

dat_race <- na.omit(dat_raw_com) %>% select(group, race, race_num) %>% group_by(group, race) %>% summarise(value = n()/unique(race_num)) %>% ungroup() %>%  mutate(individual = c("race"),   observation = race) %>% select(individual, group, observation, value)
dat_educ <- na.omit(dat_raw_com) %>% select(group, educ, educ_num) %>% group_by(group, educ) %>% summarise(value = n()/unique(educ_num)) %>% ungroup() %>% mutate(individual = c("educ"), observation = educ) %>%  select(individual, group, observation, value)
dat_income <- na.omit(dat_raw_com) %>% select(group, income, income_num) %>% group_by(group, income) %>% summarise(value = n()/unique(income_num)) %>% ungroup() %>% mutate(individual = c("income"), observation = income) %>% select(individual, group, observation, value)
data_bar <- rbind.data.frame(dat_race, dat_educ, dat_income) %>% arrange(group, individual, observation)
```

###population_disease
```{r}
dat_raw_com %>% group_by(group) %>% summarise(population = n()) %>% mutate(groupname = c("bad mental status",  "heart attack", "angina or coronary heart disease", "stroke", "asthma")) %>% select(groupname, population)
```

Among the interviewed population, the number of people suffering from potential mental illness or asthma is much larger than the number of people with heart diseases. 


###bar_plot1
```{r}
# library
data_bar_heart <- data_bar %>% filter(!group %in%  c("A","E"))

library(tidyverse)
library(viridis)
 
# Set a number of 'empty bar' to add at the end of each group
empty_bar=2
nObsType=nlevels(as.factor(data_bar_heart$observation))
to_add = data.frame( matrix(NA, empty_bar*nlevels(as.factor(data_bar_heart$group))*nObsType, ncol(data_bar_heart)) )
colnames(to_add) = colnames(data_bar_heart)
to_add$group=rep(levels(as.factor(data_bar_heart$group)), each=empty_bar*nObsType )
data_bar_heart=rbind(data_bar_heart, to_add)
data_bar_heart=data_bar_heart%>% arrange(group, individual)
data_bar_heart$id=c(rep( seq(1, 15) , c(rep(c(3,4,4,4,4),3))))

# Get the name and the y position of each label
label_data= data_bar_heart %>% group_by(id, individual) %>% summarize(tot=sum(value))
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
 
# prepare a data frame for base lines
base_data=data_bar_heart %>% 
  group_by(group) %>% 
  summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))
 
# prepare a data frame for grid (scales)
grid_data = base_data
grid_data$end = grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start = grid_data$start - 1
grid_data=grid_data[-1,]
 
# Make the plot
p = ggplot(data_bar_heart) +      
  
  # Add the stacked bar
  geom_bar(aes(x=as.factor(id), y=value, fill=observation), stat="identity", alpha=0.5) +
  scale_fill_viridis(discrete=TRUE) +
  
  # Add a val=100/200/300/400 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 0, xend = start, yend = 0), colour = "grey", alpha=1, size=1 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0.08, xend = start, yend = 0.08), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0.16, xend = start, yend = 0.16), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0.24, xend = start, yend = 0.24), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0.32, xend = start, yend = 0.32), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  
  # Add text showing the value of each 100/75/50/25 lines
  annotate("text", x = rep(max(data_bar_heart$id),5), y = c(0, 0.08, 0.16, 0.24, 0.32), label = c("0", "8%", "16%", "24%", "32%") , color="grey", size=6 , angle=0, fontface="bold", hjust=1) +
  
  ylim(-0.16,max(label_data$tot+0.08, na.rm=T)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() +
  
  # Add labels on top of each bar
  geom_text(data=label_data, aes(x=id, y=tot+0.02, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=5, angle= label_data$angle, inherit.aes = FALSE ) +
  
  # Add base line information
  geom_segment(data=base_data, aes(x = start, y = -0.0001, xend = end, yend = -0.0001), colour = "black", alpha=0.8, size=0.6 , inherit.aes = FALSE )  +
  geom_text(data=base_data, aes(x = title, y = -0.02, label=group), hjust=c(1,1,0), colour = "black", alpha=0.8, size=4, fontface="bold", inherit.aes = FALSE)
p
```



###bar_plot2
```{r}
data_bar_disease <- data_bar %>% filter(group %in% c("A", "B", "E"))
# library
library(tidyverse)
library(viridis)

# Set a number of 'empty bar' to add at the end of each group
empty_bar=2
nObsType=nlevels(as.factor(data_bar_disease$observation))
to_add = data.frame( matrix(NA, empty_bar*nlevels(as.factor(data_bar_disease$group))*nObsType, ncol(data_bar_disease)))
colnames(to_add) = colnames(data_bar_disease)
to_add$group=rep(levels(as.factor(data_bar_disease$group)), each=empty_bar*nObsType )
data_bar_disease=rbind(data_bar_disease, to_add)
data_bar_disease=data_bar_disease %>% arrange(group, individual)
data_bar_disease$id=c(rep( seq(1, 15) , c(rep(c(3,4,4,4,4),3))))
 
# Get the name and the y position of each label
label_data= data_bar_disease %>% group_by(id, individual) %>% summarize(tot=sum(value))
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
 
# prepare a data frame for base lines
base_data=data_bar_disease %>% 
  group_by(group) %>% 
  summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))
 
# prepare a data frame for grid (scales)
grid_data = base_data
grid_data$end = grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start = grid_data$start - 1
grid_data=grid_data[-1,]
 
# Make the plot
p = ggplot(data_bar_disease) +      
  
  # Add the stacked bar
  geom_bar(aes(x=as.factor(id), y=value, fill=observation), stat="identity", alpha=0.5) +
  scale_fill_viridis(discrete=TRUE) +
  
  # Add a val=1.6/1.2/0.8/0.4 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 0, xend = start, yend = 0), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0.4, xend = start, yend = 0.4), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0.8, xend = start, yend = 0.8), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 1.2, xend = start, yend = 1.2), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 1.6, xend = start, yend = 1.6), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  
  # Add text showing the value of each 100/75/50/25 lines
  annotate("text", x = rep(max(data_bar_disease$id),5), y = c(0, 0.4, 0.8, 1.2, 1.6), label = c("0", "0.4", "0.8", "1.2", "1.6") , color="black", size=6 , angle=0, fontface="bold", hjust=1) +
  
  ylim(-0.13,max(label_data$tot+0.4, na.rm=T)) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  
  # Add labels on top of each bar
  geom_text(data=label_data, aes(x=id, y=-0.12, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=4, inherit.aes = FALSE ) +
  
  # Add base line information
  geom_segment(data=base_data, aes(x = start, y = -0.0001, xend = end, yend = -0.0001), colour = "black", alpha=0.8, size=0.6 , inherit.aes = FALSE )  +
  geom_text(data=base_data, aes(x = title, y = -0.05, label=group), colour = "black", alpha=0.8, size=4, fontface="bold", inherit.aes = FALSE)
p
```

The two barplots above are used to discribe different proportions of patients with a certain disease in a group of people with a same socioeconomic feature. For example, let's look at the first bar of group A in the left plot. The length of green bar means the proportion of people with bad mental status (feel not good more than 10 days in a month) in a group of people with graduate level of education. Considering with other two bars with different color, it is not hard to come to a conclusion that the proportion of people with bad mental status in high-income group is lower than the proportion of patients in low-income group.
Socioeconomic patterns of the five diseases are very similar：
The proportion of patients in high-income group is lower than the proportion of patients in low-income group;
The proportion of patients in high-educated groups is lower than the proportion of patients in poor-educated groups;
The proportion of patients in Asian people is the lowest, while the proportion of patients in black people is the highest;




#More data 
```{r}
quantile(dat$menthlth, 0.5, na.rm = TRUE)
quantile(dat$heartattack, 0.5, na.rm = TRUE)
quantile(dat$ACheartdis, 0.5, na.rm = TRUE)
quantile(dat$stroke, 0.5, na.rm = TRUE)
quantile(dat$asthma, 0.5, na.rm = TRUE)
dat_class <- dat%>% filter(!is.na(menthlth) & !is.na(heartattack) & 
                             !is.na(ACheartdis) & !is.na(stroke) & !is.na(asthma)) %>%
  mutate(menthlth_class = ifelse(menthlth >= 10, 1, 0)) %>%
  mutate(heartattack_class = ifelse(heartattack >= 0.06, 1, 0)) %>%
  mutate(ACheartdis_class = ifelse(ACheartdis >= 0.06, 1, 0)) %>%
  mutate(stroke_class = ifelse(stroke >= 0.04, 1, 0)) %>%
  mutate(asthma_class = ifelse(asthma >= 0.12, 1, 0)) 
```






#lifestyple VS mental_class
```{r}
library(car)
library(RColorBrewer)
train_set <- read.csv("train.csv")
test_set <- read.csv("test.csv")
# Make the plot
my_colors <- brewer.pal(nlevels(as.factor(train_set$mental_class)), "Set2")
scatterplotMatrix(~smoke+drink+exercise|mental_class, data=train_set , reg.line="" , smoother="", col=my_colors , smoother.args=list(col="grey") , cex=0.5 , pch=c(15,16,17), main="The Impact of Lifestyle on Mental Status")
 

```

Correlation is not causation. However,  the exploration of correlations between lifestyle and mental status would probably help us remain in a relatively better mental status. So, what could we learn from the plot above?
The dataset to make this plot is based on the US county level with "smoke" representing the proportion of people ever diagnosed with smoking in a county, "drink" representing the proportion of people ever diagnosed with drinking and "exercise" representing the proportion of people exercised in the past 30 days.
Not surprising, counties where majorities of people have a bad mental status (green) have a larger smoking group and smaller exercise group. However, a larger drinking group in such counties is inconsistent with our expectation. 
So why people what to drink? Do you think people tend to immersed in sadness and alcohol? Probably not!  They may simply drink trouble away!



#plot for missing data
```{r}
library(dplyr)
library(wakefield)
library(Amelia)
missmap(dat, main = "Missingness Map", col = c("white", "darkgreen"), rank.order = TRUE)
```

This missingness map shows where missingness occurs in the dataset.  The order of the variables along the the x-axis is sorted by the percent missing. As shown in the plot, there are two many missing information in terms of heart attack, angina or coronary heart disease and stroke. Relatively, if we want to continue our analysis using regression or classification methods, menthlth and asthma should be better choices. 
As for our group, we are more interested in what socioeconomic featrures and environmental exposures would impact people's mental status and what are their large are their influence respectively.
