---
title: "prepare data set"
author: "Chuying Ma"
date: "11/19/2017"
output: html_document
---

BST 260 Final Project
Group: Chuying Ma, Jingjing Tang, Jie Yin, Xuewei Zhang

This Rmd file is for data cleaning and aggregating BRFSS data into county level for the year 2007


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library we need:
```{r}
library(SASxport)
library(maps)
library(dplyr)
library(tidyverse)
library(tidyr)
library(SASxport)

data_07 = read.xport("~/Desktop/Harvard/fall 2017/bst260/project/CDBRFS07.XPT")
```

try to aggregate for the data in 2000


select useful columns:
```{r}
dat_col_07 = data_07 %>%
  select(X.STATE,CTYCODE,MSCODE,IMONTH,MENTHLTH,CVDINFR4,CVDCRHD4,CVDSTRK3,ASTHMA2,AGE,SEX,ORACE2,EDUCA,INCOME2,MARITAL,EMPLOY,HLTHPLAN,EXERANY2,X.RFSMOK3,DRNKANY4,X.BMI4CAT,X.BMI4,GENHLTH,QLACTLM2)


```

create fips and create year variable:
```{r}
dat_col_07 = dat_col_07 %>%
  filter(!is.na(CTYCODE) & !is.na(X.STATE)) %>%
  mutate(fips = 1000* as.numeric(X.STATE) + as.numeric(CTYCODE)) %>%
  mutate(year = 2007) %>%
  select(fips, X.STATE,CTYCODE,MSCODE,IMONTH,MENTHLTH,CVDINFR4,CVDCRHD4,CVDSTRK3,ASTHMA2,AGE,SEX,ORACE2,EDUCA,INCOME2,MARITAL,EMPLOY,HLTHPLAN,EXERANY2,X.RFSMOK3,DRNKANY4,X.BMI4CAT,X.BMI4,GENHLTH,QLACTLM2)

```

start aggregate into county level:
1.start with menthlth
```{r}
menthlth = dat_col_07 %>%
  select(fips,MENTHLTH) %>%
  filter(!is.na(MENTHLTH) & MENTHLTH != 77 & MENTHLTH != 88 & MENTHLTH != 99) %>%
  group_by(fips) %>%
  dplyr::summarise(menthlth = mean(MENTHLTH))
```

2.hearattack
```{r}
count_fips = dat_col_07 %>%
  select(fips,CVDINFR4) %>%
  filter(!is.na(CVDINFR4) & CVDINFR4 != 7 & CVDINFR4 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_h = dat_col_07 %>%
  group_by(fips,CVDINFR4) %>%
  dplyr::summarise(n = n())

tab_heart = count_h %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(hearattack = n/countF) %>%
  filter(!is.na(CVDINFR4) & CVDINFR4 == 1) %>%
  select(fips,hearattack)

```

3.ACheartdis
```{r}
count_fips = dat_col_07 %>%
  select(fips,CVDCRHD4) %>%
  filter(!is.na(CVDCRHD4) & CVDCRHD4 !=7 & CVDCRHD4 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_a = dat_col_07 %>%
  group_by(fips,CVDCRHD4) %>%
  dplyr::summarise(n = n())


tab_ac = count_a %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(ACheartdis = n/countF) %>%
  filter(!is.na(CVDCRHD4) & CVDCRHD4 == 1) %>%
  select(fips,ACheartdis)
```

4. stroke
```{r}
count_fips = dat_col_07 %>%
  select(fips,CVDSTRK3) %>%
  filter(!is.na(CVDSTRK3) & CVDSTRK3 != 7 & CVDSTRK3 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_s = dat_col_07 %>%
  group_by(fips,CVDSTRK3) %>%
  dplyr::summarise(n = n())


tab_st = count_s %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(stroke = n/countF) %>%
  filter(!is.na(CVDSTRK3) & CVDSTRK3 == 1) %>%
  select(fips,stroke)
```

5. asthma
```{r}
count_fips = dat_col_07 %>%
  select(fips,ASTHMA2) %>%
  filter(!is.na(ASTHMA2) & ASTHMA2 != 7 & ASTHMA2 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_as = dat_col_07 %>%
  group_by(fips,ASTHMA2) %>%
  dplyr::summarise(n = n())

tab_as = count_as %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(asthma = n/countF) %>%
  filter(!is.na(ASTHMA2) & ASTHMA2 == 1) %>%
  select(fips,asthma)
```

6. age
```{r}
age = dat_col_07 %>%
  select(fips,AGE) %>%
  filter(!is.na(AGE) & AGE != 7, AGE != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(age = mean(AGE))
```

7.sex
```{r}
count_fips = dat_col_07 %>%
  select(fips,SEX) %>%
  filter(!is.na(SEX)) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_sex = dat_col_07 %>%
  group_by(fips,SEX) %>%
  dplyr::summarise(n = n())

tab_sex = count_sex %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(sex = n/countF) %>%
  filter(!is.na(SEX) & SEX == 2) %>%
  select(fips,sex)
```

8.race
```{r}
count_fips = dat_col_07 %>%
  select(fips,ORACE2) %>%
  filter(!is.na(ORACE2) & ORACE2 != 7 & ORACE2 != 8 & ORACE2 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_race = dat_col_07 %>%
  group_by(fips,ORACE2) %>%
  dplyr::summarise(n = n())

white = count_race %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(white = n/countF) %>%
  filter(!is.na(ORACE2) & ORACE2 == 1) %>%
  select(fips,white)

black = count_race %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(black = n/countF) %>%
  filter(!is.na(ORACE2) & ORACE2 == 2) %>%
  select(fips,black)

asian = count_race %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(asian = n/countF) %>%
  filter(!is.na(ORACE2) & ORACE2 == 3) %>%
  select(fips,asian)

other_race = count_race %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(race1 = n/countF) %>%
  filter(!is.na(ORACE2) & ORACE2 %in% c(4,5,6)) %>%
  group_by(fips) %>%
  dplyr::summarise(race_other = sum(race1)) %>%
  select(fips,race_other)
```

9. educ
```{r}
count_fips = dat_col_07 %>%
  select(fips,EDUCA) %>%
  filter(!is.na(EDUCA) & EDUCA != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_edu = dat_col_07 %>%
  group_by(fips,EDUCA) %>%
  dplyr::summarise(n = n())

tab_educ1 = count_edu %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(educ = n/countF) %>%
  filter(!is.na(EDUCA) & EDUCA %in% c(1,2)) %>%
  group_by(fips) %>%
  dplyr::summarise(educ1 = sum(educ)) %>%
  select(fips,educ1)

tab_educ2 = count_edu %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(educ = n/countF) %>%
  filter(!is.na(EDUCA) & EDUCA %in% c(3,4)) %>%
  group_by(fips) %>%
  dplyr::summarise(educ2 = sum(educ)) %>%
  select(fips,educ2)

tab_educ3 = count_edu %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(educ = n/countF) %>%
  filter(!is.na(EDUCA) & EDUCA %in% c(5,6)) %>%
  group_by(fips) %>%
  dplyr::summarise(educ3 = sum(educ)) %>%
  select(fips,educ3)
```

10.income
```{r}
count_fips = dat_col_07 %>%
  select(fips,INCOME2) %>%
  filter(!is.na(INCOME2) & INCOME2 != 77 & INCOME2 != 99) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_inc = dat_col_07 %>%
  group_by(fips,INCOME2) %>%
  dplyr::summarise(n = n())

tab_inc1 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income1 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 01) %>%
  select(fips,income1)

tab_inc2 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income2 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 02) %>%
  select(fips,income2)

tab_inc3 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income3 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 03) %>%
  select(fips,income3)

tab_inc4 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income4 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 04) %>%
  select(fips,income4)

tab_inc5 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income5 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 05) %>%
  select(fips,income5)

tab_inc6 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income6 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 06) %>%
  select(fips,income6)

tab_inc7 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income7 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 07) %>%
  select(fips,income7)

tab_inc8 = count_inc %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(income8 = n/countF) %>%
  filter(!is.na(INCOME2) & INCOME2 == 08) %>%
  select(fips,income8)
```

11.marital
```{r}
count_fips = dat_col_07 %>%
  select(fips,MARITAL) %>%
  filter(!is.na(MARITAL) & MARITAL != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_mar = dat_col_07 %>%
  group_by(fips,MARITAL) %>%
  dplyr::summarise(n = n())

tab_married = count_mar %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(married = n/countF) %>%
  filter(!is.na(MARITAL) & MARITAL == 1) %>%
  select(fips,married)

tab_unmar = count_mar %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(unmarried1 = n/countF) %>%
  filter(!is.na(MARITAL) & MARITAL %in% c(2,3,4,5,6)) %>%
  group_by(fips) %>%
  dplyr::summarise(unmarried = sum(unmarried1)) %>%
  select(fips,unmarried)
```

12.employ
```{r}
count_fips = dat_col_07 %>%
  select(fips,EMPLOY) %>%
  filter(!is.na(EMPLOY) & EMPLOY != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_emp = dat_col_07 %>%
  group_by(fips,EMPLOY) %>%
  dplyr::summarise(n = n())

tab_employed = count_emp %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(employed1 = n/countF) %>%
  filter(!is.na(EMPLOY) & EMPLOY %in% c(1,2)) %>%
  group_by(fips) %>%
  dplyr::summarise(employed = sum(employed1)) %>%
  select(fips,employed)
  
tab_out = count_emp %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(employed1 = n/countF) %>%
  filter(!is.na(EMPLOY) & EMPLOY %in% c(3,4)) %>%
  group_by(fips) %>%
  dplyr::summarise(out_of_work = sum(employed1)) %>%
  select(fips,out_of_work)
  
tab_home = count_emp %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(employed1 = n/countF) %>%
  filter(!is.na(EMPLOY) & EMPLOY %in% c(5)) %>%
  group_by(fips) %>%
  dplyr::summarise(homemaker = sum(employed1)) %>%
  select(fips,homemaker)
  
tab_stu = count_emp %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(employed1 = n/countF) %>%
  filter(!is.na(EMPLOY) & EMPLOY %in% c(6)) %>%
  group_by(fips) %>%
  dplyr::summarise(student = sum(employed1)) %>%
  select(fips,student)
  
tab_emp_other = count_emp %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(employed1 = n/countF) %>%
  filter(!is.na(EMPLOY) & EMPLOY %in% c(7,8)) %>%
  group_by(fips) %>%
  dplyr::summarise(employ_other = sum(employed1)) %>%
  select(fips,employ_other)
```

13. hlthplan
```{r}
count_fips = dat_col_07 %>%
  select(fips,HLTHPLAN) %>%
  filter(!is.na(HLTHPLAN) & HLTHPLAN != 7 & HLTHPLAN != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_plan = dat_col_07 %>%
  group_by(fips,HLTHPLAN) %>%
  dplyr::summarise(n = n())

tab_plan = count_plan %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(hlthplan = n/countF) %>%
  filter(!is.na(HLTHPLAN) & HLTHPLAN == 1) %>%
  select(fips,hlthplan)
```

14.exercise
```{r}
count_fips = dat_col_07 %>%
  select(fips,EXERANY2) %>%
  filter(!is.na(EXERANY2) & EXERANY2 != 7 & EXERANY2 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_ex = dat_col_07 %>%
  group_by(fips,EXERANY2) %>%
  dplyr::summarise(n = n())

tab_ex = count_ex %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(exercise = n/countF) %>%
  filter(!is.na(EXERANY2) & EXERANY2 == 1) %>%
  select(fips,exercise)
```

15.smoke
```{r}
count_fips = dat_col_07 %>%
  select(fips,X.RFSMOK3) %>%
  filter(!is.na(X.RFSMOK3) & X.RFSMOK3 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_sm = dat_col_07 %>%
  group_by(fips,X.RFSMOK3) %>%
  dplyr::summarise(n = n())

tab_sm = count_sm %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(smoke = n/countF) %>%
  filter(!is.na(X.RFSMOK3) & X.RFSMOK3 == 2) %>%
  select(fips,smoke)
```

16.drink
```{r}
count_fips = dat_col_07 %>%
  select(fips,DRNKANY4) %>%
  filter(!is.na(DRNKANY4) & DRNKANY4 != 7 & DRNKANY4 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_drk = dat_col_07 %>%
  group_by(fips,DRNKANY4) %>%
  dplyr::summarise(n = n())

tab_drk = count_drk %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(drink = n/countF) %>%
  filter(!is.na(DRNKANY4) & DRNKANY4 == 1) %>%
  select(fips,drink)
```

17.bmi_cat
```{r}
count_fips = dat_col_07 %>%
  select(fips,X.BMI4CAT) %>%
  filter(!is.na(X.BMI4CAT) & X.BMI4CAT != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_bmicat = dat_col_07 %>%
  group_by(fips,X.BMI4CAT) %>%
  dplyr::summarise(n = n())

tab_bmicat1 = count_bmicat %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(bmi_cat1 = n/countF) %>%
  filter(!is.na(X.BMI4CAT) & X.BMI4CAT == 1) %>%
  select(fips,bmi_cat1)

tab_bmicat2 = count_bmicat %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(bmi_cat2 = n/countF) %>%
  filter(!is.na(X.BMI4CAT) & X.BMI4CAT == 2) %>%
  select(fips,bmi_cat2)

tab_bmicat3 = count_bmicat %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(bmi_cat3 = n/countF) %>%
  filter(!is.na(X.BMI4CAT) & X.BMI4CAT == 3) %>%
  select(fips,bmi_cat3)
```

18.bmi_cts
```{r}
tab_bmicts = dat_col_07 %>%
  select(fips,X.BMI4) %>%
  filter(X.BMI4 != 9999) %>%
  group_by(fips) %>%
  dplyr::summarise(bmi_cts = mean(X.BMI4))
```


19.genhlth
```{r}
count_fips = dat_col_07 %>%
  select(fips,GENHLTH) %>%
  filter(!is.na(GENHLTH) & GENHLTH != 7  & GENHLTH != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_genh = dat_col_07 %>%
  group_by(fips,GENHLTH) %>%
  dplyr::summarise(n = n())

tab_genh1 = count_genh %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(genhlth_ex = n/countF) %>%
  filter(!is.na(GENHLTH) & GENHLTH == 1) %>%
  select(fips,genhlth_ex)

tab_genh2 = count_genh %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(genhlth_vg = n/countF) %>%
  filter(!is.na(GENHLTH) & GENHLTH == 2) %>%
  select(fips,genhlth_vg)

tab_genh3 = count_genh %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(genhlth_go = n/countF) %>%
  filter(!is.na(GENHLTH) & GENHLTH == 3) %>%
  select(fips,genhlth_go)

tab_genh4 = count_genh %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(genhlth_fa = n/countF) %>%
  filter(!is.na(GENHLTH) & GENHLTH == 4) %>%
  select(fips,genhlth_fa)

tab_genh5 = count_genh %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(genhlth_po = n/countF) %>%
  filter(!is.na(GENHLTH) & GENHLTH == 5) %>%
  select(fips,genhlth_po)

```

20.qlactlm
```{r}
count_fips = dat_col_07 %>%
  select(fips,QLACTLM2) %>%
  filter(!is.na(QLACTLM2) & QLACTLM2 != 7 & QLACTLM2 != 9) %>%
  group_by(fips) %>%
  dplyr::summarise(countF = n())
  
count_ql = dat_col_07 %>%
  group_by(fips,QLACTLM2) %>%
  dplyr::summarise(n = n())

tab_ql = count_ql %>% 
  left_join(count_fips,by = c('fips' = 'fips')) %>%
  mutate(qlactlm = n/countF) %>%
  filter(!is.na(QLACTLM2) & QLACTLM2 == 1) %>%
  select(fips,qlactlm)
```

left join all sub-tables:
```{r,eval=FALSE}
tab_first = dat_col_07 %>%
  mutate(year = 2007) %>%
  select(fips,state = X.STATE,county = CTYCODE, metro = MSCODE,year)

tab_first = unique(tab_first)
```

```{r,eval=FALSE}
#var7
final_00 = tab_first %>%
  left_join(menthlth,by= c('fips' = 'fips'))

#var8
final_00 = final_00 %>%
  left_join(tab_heart,by= c('fips' = 'fips'))

#var9
final_00 = final_00 %>%
  left_join(tab_ac,by= c('fips' = 'fips'))

#var10
final_00 = final_00 %>%
  left_join(tab_st,by= c('fips' = 'fips'))

#var11
final_00 = final_00 %>%
  left_join(tab_as,by= c('fips' = 'fips'))

#var12
final_00 = final_00 %>%
  left_join(age,by= c('fips' = 'fips'))

#var13
final_00 = final_00 %>%
  left_join(tab_sex,by= c('fips' = 'fips'))

#var14
final_00 = final_00 %>%
  left_join(white,by= c('fips' = 'fips'))

#var15
final_00 = final_00 %>%
  left_join(black,by= c('fips' = 'fips'))

#var16
final_00 = final_00 %>%
  left_join(asian,by= c('fips' = 'fips'))

#var17
final_00 = final_00 %>%
  left_join(other_race,by= c('fips' = 'fips'))

#var18 ~20
final_00 = final_00 %>%
  left_join(tab_educ1,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_educ2,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_educ3,by= c('fips' = 'fips'))

#var20~27
final_00 = final_00 %>%
  left_join(tab_inc1,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc2,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc3,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc4,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc5,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc6,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc7,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_inc8,by= c('fips' = 'fips'))

#var28 & 29
final_00 = final_00 %>%
  left_join(tab_married,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_unmar,by= c('fips' = 'fips'))

#var 30~34
final_00 = final_00 %>%
  left_join(tab_employed,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_out,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_home,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_stu,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_emp_other,by= c('fips' = 'fips'))

#var35
final_00 = final_00 %>%
  left_join(tab_plan,by= c('fips' = 'fips'))

#var36
final_00 = final_00 %>%
  left_join(tab_ex,by= c('fips' = 'fips'))

#var 37
final_00 = final_00 %>%
  left_join(tab_sm,by= c('fips' = 'fips'))

#var38
final_00 = final_00 %>%
  left_join(tab_drk,by= c('fips' = 'fips'))

#var 39~41
final_00 = final_00 %>%
  left_join(tab_bmicat1,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_bmicat2,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_bmicat3,by= c('fips' = 'fips'))

#var 42
final_00 = final_00 %>%
  left_join(tab_bmicts,by= c('fips' = 'fips'))

#var43 ~ 47
final_00 = final_00 %>%
  left_join(tab_genh1,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_genh2,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_genh3,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_genh4,by= c('fips' = 'fips'))
final_00 = final_00 %>%
  left_join(tab_genh5,by= c('fips' = 'fips'))

#var48
final_00 = final_00 %>%
  left_join(tab_ql,by= c('fips' = 'fips'))
```

save as csv:
```{r,eval=FALSE}
write.csv(final_00,file = "2007_data.csv",row.names = F)
```

more steps are needed to finalize the model since I didn't consider the presense of "777" and "999" in the CTYCODE
```{r,eval=FALSE}
data_2007 = read_csv("2007_data.csv")
data_2007 = data_2007 %>%
  filter(county != 777 & county != 999) %>%
  mutate(bmi_cts = bmi_cts/100) %>%
  select(-metro)

data_2007 = unique(data_2007)

write_csv(data_2007,"~/Desktop/Harvard/fall 2017/bst260/project/data_2007.csv")
```